## Backgrownd

For this assignment, I build a predictive model to determine whether a particular form of exercise (barbell lifting) 
is performed correctly, using accelerometer data.

## Loading of the packages
The first part is the declaration of the package which will be used. 

Note : to be reproductible, I also set the seed value.

```{r}
library(caret)
library(randomForest)
library(Hmisc)

set.seed(1234)
```
## Reading of the Data

We begin by reading in the training and testing datasets, 
assigning missing values to entries that are currently 'NA' or blank.
(commands are commented to limit the output size. You can run it deleting the "#" ) 

```{r}
#getwd()
setwd("C:/PredictiveAnalysis/R")

train <- read.csv("./data/pml-training.csv", header = TRUE, na.strings = c("NA", ""))
test <- read.csv("./data/pml-testing.csv", header = TRUE, na.strings = c("NA", ""))

```
#dim(test)
#names(train)

dim(train)

We see that the training set consists of 19622 observations of 160 variables
```{r}
sum(complete.cases(train))
#head(train)
```


```{r}






